### 2024-04-18 paper/paper.qmd
commit

---
title: "Social Housing in Toronto: Investigating Status"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Results

The dataset "Centralized Waitlist for Social Housing" provides detailed information on the activities related to the social housing waitlist in Toronto for various quarters starting from Q1 2019. It includes data on new/reactivated applications, the number of households housed, inactive/cancelled applications, and the total active waiting list. Additionally, it breaks down the waiting list by household type (e.g., no dependents, with dependents, seniors).

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Load required packages
library(ggplot2)

# Extract relevant columns with adjusted column names
year_quarter <- raw_data$Quarter_Year
new_reactivated_applications <- raw_data$New_reactivated
housed <- raw_data$Housed

# Create a data frame
plot_data <- data.frame(Year_Quarter = year_quarter,
                        New_Reactivated_Applications = new_reactivated_applications,
                        Housed = housed)

# Convert Year_Quarter to factor to maintain order
plot_data$Year_Quarter <- factor(plot_data$Year_Quarter, levels = unique(plot_data$Year_Quarter))

# Create a line plot
ggplot(plot_data, aes(x = Year_Quarter)) +
  geom_line(aes(y = New_Reactivated_Applications, color = "New/Reactivated Applications"), size = 1.2) +
  geom_line(aes(y = Housed, color = "Housed"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Yearly New/Reactivated Applications vs. Housed",
       x = "Year Quarter",
       y = "Application Status") +
  theme_minimal() +
  theme(legend.title = element_blank())

```
The line graph illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023. The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020. Despite these fluctuations, the number of households housed each quarter consistently lags behind new entries, underscoring a substantial gap between demand and supply dynamics. This persistent shortfall indicates the ongoing challenge of meeting the housing needs of vulnerable populations within the community.

Examining waiting list trends further illuminates the complexities involved in managing social housing demand. The total active waiting list experiences fluctuations, notably declining in Q2 2019 due to a significant number of inactive or cancelled applications. This fluctuation suggests challenges in maintaining accurate waiting lists or changes in applicants' circumstances, highlighting the multifaceted nature of addressing social housing needs. Such insights underscore the importance of robust data management and adaptive policies to effectively respond to evolving housing demands.

Moreover, an analysis of the waiting list composition reveals critical insights into the demographics of those in need of social housing. A significant portion of seniors and households without dependents feature prominently in the waiting list, indicating a pressing need for housing support among these vulnerable groups. This demographic composition underscores broader social and economic challenges, including affordability issues for seniors and single individuals, necessitating targeted interventions and policy measures to address housing inequities and ensure inclusive access to adequate housing for all segments of the population.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}
The observed dynamics in social housing demand and supply, as evidenced by the fluctuating patterns in application rates and the persistent gap between new entries and households housed, underscore the complex interplay of socio-economic factors influencing housing access and affordability. Research by Desmond and Kimbro (2015) highlights the intricate relationship between housing instability and poverty, wherein marginalized communities often face challenges in accessing stable and affordable housing options. Moreover, studies by Fitzpatrick et al. (2019) emphasize the role of structural inequalities and systemic barriers in perpetuating housing disparities, particularly among vulnerable populations such as seniors and single individuals. The phenomenon of fluctuating waiting list trends and demographic compositions may be attributed to various factors, including inadequate affordable housing stock, rising housing costs, and limited access to supportive services and resources. Additionally, the complexities of managing waiting lists and addressing changing applicant circumstances further contribute to the observed variations in social housing demand. Effective policy interventions aimed at addressing housing inequities and enhancing housing affordability are crucial in mitigating these challenges and ensuring equitable access to housing for all members of society (Desmond & Kimbro, 2015; Fitzpatrick et al., 2019).

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-04-18 paper/paper.qmd
commit

@@
 ---
-title: "Social Housing in Toronto: Investigating Status"
+title: "Social Housing in Toronto: Investigating Status from Waiting Lists and Their Social Implications"
 subtitle: "My subtitle if needed"
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# Load required packages
+library(ggplot2)
+
+# Extract relevant columns with adjusted column names
+quarter/year <- raw_data$Quarter/Year
+total_active <- raw_data$Total_active_waiting_list
+seniors <- raw_data$Seniors
+
+# Create a data frame
+plot_data <- data.frame(Quarter/Year = quarter/year,
+                        Total_active_waiting_list = total_active,
+                        Seniors = seniors)
+
+# Convert Quarter/Year to factor to maintain order
+plot_data$Quarter/Year <- factor(plot_data$Quarter/Year, levels = unique(plot_data$Year/Quarter))
+
+# Create a line plot
+ggplot(plot_data, aes(x = Quarter_Year)) +
+  geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Applications"), size = 1.2) +
+  geom_line(aes(y = Seniors, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
+  scale_color_manual(values = c("blue", "red")) +
+  labs(title = "Total Active Applications vs. Seniors' Applications",
+       x = "Quarter Year",
+       y = "Application Numbers") +
+  theme_minimal() +
+  theme(legend.title = element_blank())
+
 ```
+The line plot comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
+From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period.
+However, seniors' applications appear to comprise a relatively smaller proportion of the total active applications, as indicated by the consistently lower values for seniors' applications compared to total active applications.
+This disparity suggests that while seniors constitute a significant demographic group in need of social housing support, they may not represent the majority of applicants.
+Understanding this dynamic is crucial for policymakers and housing providers to tailor their strategies and resources effectively to address the diverse needs of applicants across different age groups.
+
+Moreover, the fluctuating trends in both total active applications and seniors' applications highlight the dynamic nature of housing demand and the complex socio-economic factors influencing individuals' housing situations.
+Further analysis may be warranted to explore the underlying reasons behind these fluctuations, such as changes in demographic trends, economic conditions, or shifts in housing policies and programs.
+Overall, this graph underscores the importance of considering demographic diversity and changing trends in social housing applications when developing and implementing policies and programs aimed at addressing housing needs and promoting housing equity for all segments of the population.
 
@@
 ## Second discussion point
+
+The visualization of total active applications versus seniors' applications from Q1 2019 to Q4 2023 sheds light on the significant representation of seniors within the social housing applicant pool.
+The observation that seniors comprise at least one-third of the total applications underscores the pronounced housing needs among older adults within the community.
+Research by Morrow-Howell et al.
+(2016) emphasizes the growing demographic shift towards an aging population and the subsequent implications for housing demand, particularly among low-income seniors.
+Furthermore, studies by Burholt and Windle (2006) highlight the multifaceted challenges faced by older adults in accessing suitable housing options, including financial constraints, physical health limitations, and social isolation.
+The substantial presence of seniors within the applicant pool suggests a pressing need for targeted interventions and policy measures to address the unique housing needs of this vulnerable demographic group (Morrow-Howell et al., 2016; Burholt & Windle, 2006).
+
+Moreover, the high proportion of seniors' applications relative to the total applicant pool underscores the importance of incorporating age-specific considerations into housing policy and program development.
+Research by Butler and Robison (2001) emphasizes the significance of age-in-place initiatives and supportive housing models tailored to the needs of older adults, facilitating aging in community settings and promoting independence and well-being.
+Additionally, studies by Gonyea et al.
+(2017) highlight the benefits of integrated housing and service programs for seniors, addressing both housing and health-related needs and enhancing overall quality of life.
+The findings from the visualization underscore the imperative for comprehensive housing strategies that prioritize the housing security and dignity of seniors while addressing broader societal challenges related to aging and housing inequities (Butler & Robison, 2001; Gonyea et al., 2017).
 

### 2024-04-18 paper/paper.qmd
commit

@@
 ---
-title: "Social Housing in Toronto: Investigating Status from Waiting Lists and Their Social Implications"
-subtitle: "My subtitle if needed"
+title: "Social Housing in Toronto"
+subtitle: "Investigating Status from Waiting Lists and Their Social Implications"
 author: 
-  - First author
-  - Another author
+  - Xiaoyu (Alice) Miao
 thanks: "Code and data are available at: LINK."
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+Access to stable and affordable housing is a fundamental aspect of societal well-being, yet disparities persist in housing outcomes across different demographic groups.
+This paper investigates the dynamics of household composition in social housing applications in Toronto and its implications for housing policy and practice.
+Analyzing social housing application data from Q1 2019 to Q4 2023 within Toronto's housing landscape, the study focuses on the patterns and fluctuations in households applying for housing assistance, particularly those with and without dependents.
+The analysis highlights the diverse challenges faced by families in accessing stable and affordable housing in Toronto, emphasizing the dynamic nature of family dynamics and economic circumstances over time.
 
-The remainder of this paper is structured as follows. @sec-data....
+By recognizing the unique needs of households with dependents and understanding the socio-economic drivers of housing disparities in Toronto, policymakers and housing providers can develop targeted interventions to promote housing equity and stability for vulnerable families in the city.
+Through evidence-based policy interventions and a deeper understanding of household composition in social housing, this paper aims to contribute to the ongoing discourse on housing affordability, access, and equity in Toronto, striving towards inclusive and sustainable housing solutions for all members of Toronto's diverse community.
 
@@
 Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+
+Data Sourcing
+
+Data is sourced from Open Data Toronto.
 
@@
 
-Talk way more about it.
+The maps of Toronto showing where those social housing are spread gives the reader a more specific sense of what they look like in Toronto.
 
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Load required packages
+library(ggplot2)
+
+# Assuming your dataset is loaded into a variable named 'raw_data'
+# Extract relevant columns with adjusted column names
+year_quarter <- raw_data$Quarter_Year
+household_no_dependents <- raw_data$Household_no_dependents
+household_with_dependents <- raw_data$Household_with_dependents
+
+# Create a data frame
+plot_data <- data.frame(Year_Quarter = year_quarter,
+                        Household_no_dependents = household_no_dependents,
+                        Household_with_dependents = household_with_dependents)
+
+# Convert Year_Quarter to factor to maintain order
+plot_data$Year_Quarter <- factor(plot_data$Year_Quarter, levels = unique(plot_data$Year_Quarter))
+
+# Create a line plot
+ggplot(plot_data, aes(x = Year_Quarter)) +
+  geom_line(aes(y = Household_no_dependents, color = "Household without Dependents"), size = 1.2) +
+  geom_line(aes(y = Household_with_dependents, color = "Household with Dependents"), linetype = "dashed", size = 1.2) +
+  scale_color_manual(values = c("blue", "red")) +
+  labs(title = "Households with and without Dependents Applying for Social Housing",
+       x = "Year Quarter",
+       y = "Application Numbers") +
+  theme_minimal() +
+  theme(legend.title = element_blank())
+
+```
+
+The line plot comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
+From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period.
+Interestingly, the number of households with dependents appears to fluctuate more prominently compared to households without dependents, as indicated by the greater variability in the dashed line representing households with dependents.
+
+This variation in the number of applications from households with dependents may reflect changes in family dynamics, economic conditions, or social factors affecting household composition and housing stability.
+For instance, economic downturns or changes in family circumstances could lead to fluctuations in the number of households with dependents seeking social housing support.
+Understanding these dynamics is crucial for policymakers and housing providers to tailor their programs and services effectively to meet the diverse needs of households with different compositions.
+It highlights the importance of considering family size, structure, and specific needs when designing housing assistance programs and allocating resources to address housing insecurity and promote housing stability for vulnerable populations.
+
+Moreover, further analysis of the underlying factors driving fluctuations in applications from households with and without dependents may provide valuable insights into broader socio-economic trends and challenges affecting housing affordability, access, and stability for families in need.
+Such insights can inform evidence-based policy interventions and targeted strategies aimed at addressing housing disparities and promoting equitable access to affordable and suitable housing for all households, regardless of their composition.
+
 # Discussion
@@
 ## Third discussion point
+
+The analysis of the line plot comparing households with and without dependents applying for social housing sheds light on the complex interplay of socio-economic factors influencing housing needs and family structures.
+The observed fluctuations in applications from households with dependents suggest dynamic changes in family dynamics and economic circumstances over time.
+Research by Clampet-Lundquist et al.
+(2011) emphasizes the importance of considering household composition in housing policy and interventions, as households with dependents often face unique challenges in accessing stable and affordable housing.
+For instance, fluctuations in applications from households with dependents may reflect changes in employment opportunities, income levels, or housing affordability, which impact family decision-making regarding housing choices and stability.
+
+Moreover, studies by Murdie and Teixeira (2019) highlight the role of socio-economic factors and policy interventions in shaping housing outcomes for families, particularly those with children.
+The greater variability observed in applications from households with dependents compared to those without dependents underscores the diverse and evolving needs of families in the social housing system.
+Understanding these dynamics is essential for policymakers and housing providers to develop targeted interventions and support services that address the specific needs of families with dependents and promote housing equity for all members of society.
+
+In summary, the analysis of household compositions in social housing applications offers valuable insights into the nuanced challenges and vulnerabilities faced by families in accessing stable and affordable housing.
+By recognizing the unique needs of households with dependents and implementing evidence-based policy interventions, policymakers can work towards reducing housing disparities and promoting housing stability for vulnerable families in need (Clampet-Lundquist et al., 2011; Murdie & Teixeira, 2019).
 

### 2024-04-21 paper/paper.qmd
commit

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper investigates household composition dynamics in social housing applications in Toronto from Q1 2019 to Q4 2023.
+Analyzing application data, the study examines patterns in households applying for housing assistance, focusing on those with and without dependents.
+Findings reveal fluctuating trends, highlighting the challenges faced by families in accessing stable and affordable housing in Toronto.
+By understanding these dynamics, policymakers can develop targeted interventions to promote housing equity and stability for vulnerable families in the city, contributing to ongoing efforts to address housing affordability and access issues in Toronto."
 format: pdf
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
 Data Sourcing
@@
 Data is sourced from Open Data Toronto.
+
+Data Cleaning
 
@@
 The observed dynamics in social housing demand and supply, as evidenced by the fluctuating patterns in application rates and the persistent gap between new entries and households housed, underscore the complex interplay of socio-economic factors influencing housing access and affordability.
-Research by Desmond and Kimbro (2015) highlights the intricate relationship between housing instability and poverty, wherein marginalized communities often face challenges in accessing stable and affordable housing options.
-Moreover, studies by Fitzpatrick et al.
-(2019) emphasize the role of structural inequalities and systemic barriers in perpetuating housing disparities, particularly among vulnerable populations such as seniors and single individuals.
+Research by Desmond and @desmond2015 highlights the intricate relationship between housing instability and poverty, wherein marginalized communities often face challenges in accessing stable and affordable housing options.
+Moreover, studies by @fitzpatrick2019 emphasize the role of structural inequalities and systemic barriers in perpetuating housing disparities, particularly among vulnerable populations such as seniors and single individuals.
 The phenomenon of fluctuating waiting list trends and demographic compositions may be attributed to various factors, including inadequate affordable housing stock, rising housing costs, and limited access to supportive services and resources.
 Additionally, the complexities of managing waiting lists and addressing changing applicant circumstances further contribute to the observed variations in social housing demand.
-Effective policy interventions aimed at addressing housing inequities and enhancing housing affordability are crucial in mitigating these challenges and ensuring equitable access to housing for all members of society (Desmond & Kimbro, 2015; Fitzpatrick et al., 2019).
+Effective policy interventions aimed at addressing housing inequities and enhancing housing affordability are crucial in mitigating these challenges and ensuring equitable access to housing for all members of society @desmond2015, @fitzpatrick2019.
 
@@
 The observation that seniors comprise at least one-third of the total applications underscores the pronounced housing needs among older adults within the community.
-Research by Morrow-Howell et al.
-(2016) emphasizes the growing demographic shift towards an aging population and the subsequent implications for housing demand, particularly among low-income seniors.
-Furthermore, studies by Burholt and Windle (2006) highlight the multifaceted challenges faced by older adults in accessing suitable housing options, including financial constraints, physical health limitations, and social isolation.
-The substantial presence of seniors within the applicant pool suggests a pressing need for targeted interventions and policy measures to address the unique housing needs of this vulnerable demographic group (Morrow-Howell et al., 2016; Burholt & Windle, 2006).
+Research by @morrow2016 emphasizes the growing demographic shift towards an aging population and the subsequent implications for housing demand, particularly among low-income seniors.
+Furthermore, studies by @burholt2006 highlight the multifaceted challenges faced by older adults in accessing suitable housing options, including financial constraints, physical health limitations, and social isolation.
+The substantial presence of seniors within the applicant pool suggests a pressing need for targeted interventions and policy measures to address the unique housing needs of this vulnerable demographic group @morrow2016; @burholt2006.
 
 Moreover, the high proportion of seniors' applications relative to the total applicant pool underscores the importance of incorporating age-specific considerations into housing policy and program development.
-Research by Butler and Robison (2001) emphasizes the significance of age-in-place initiatives and supportive housing models tailored to the needs of older adults, facilitating aging in community settings and promoting independence and well-being.
-Additionally, studies by Gonyea et al.
-(2017) highlight the benefits of integrated housing and service programs for seniors, addressing both housing and health-related needs and enhancing overall quality of life.
-The findings from the visualization underscore the imperative for comprehensive housing strategies that prioritize the housing security and dignity of seniors while addressing broader societal challenges related to aging and housing inequities (Butler & Robison, 2001; Gonyea et al., 2017).
+Research by @butler2001 emphasizes the significance of age-in-place initiatives and supportive housing models tailored to the needs of older adults, facilitating aging in community settings and promoting independence and well-being.
+Additionally, studies by @gonyea2017 highlight the benefits of integrated housing and service programs for seniors, addressing both housing and health-related needs and enhancing overall quality of life.
+The findings from the visualization underscore the imperative for comprehensive housing strategies that prioritize the housing security and dignity of seniors while addressing broader societal challenges related to aging and housing inequities @butler2001; @gonyea2017.
 
@@
 The observed fluctuations in applications from households with dependents suggest dynamic changes in family dynamics and economic circumstances over time.
-Research by Clampet-Lundquist et al.
-(2011) emphasizes the importance of considering household composition in housing policy and interventions, as households with dependents often face unique challenges in accessing stable and affordable housing.
+Research by @clampet2011 emphasizes the importance of considering household composition in housing policy and interventions, as households with dependents often face unique challenges in accessing stable and affordable housing.
 For instance, fluctuations in applications from households with dependents may reflect changes in employment opportunities, income levels, or housing affordability, which impact family decision-making regarding housing choices and stability.
 
-Moreover, studies by Murdie and Teixeira (2019) highlight the role of socio-economic factors and policy interventions in shaping housing outcomes for families, particularly those with children.
+Moreover, studies by @murdie2019 highlight the role of socio-economic factors and policy interventions in shaping housing outcomes for families, particularly those with children.
 The greater variability observed in applications from households with dependents compared to those without dependents underscores the diverse and evolving needs of families in the social housing system.
@@
 In summary, the analysis of household compositions in social housing applications offers valuable insights into the nuanced challenges and vulnerabilities faced by families in accessing stable and affordable housing.
-By recognizing the unique needs of households with dependents and implementing evidence-based policy interventions, policymakers can work towards reducing housing disparities and promoting housing stability for vulnerable families in need (Clampet-Lundquist et al., 2011; Murdie & Teixeira, 2019).
+By recognizing the unique needs of households with dependents and implementing evidence-based policy interventions, policymakers can work towards reducing housing disparities and promoting housing stability for vulnerable families in need @clampet2011; @murdie2019.
 

### 2024-04-21 scripts/02-data_cleaning.R

@@ -1,44 +1,24 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw plane data 
+# Author: Xiaoyu Miao
+# Date: 17 April 2024
+# Contact: xy.miao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+#Read data
+raw_data <- read_csv("data/raw_data/Centralized Waitlist for Social Housing.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+# Get the index of column J
+start_index <- which(colnames(raw_data) == "Total")
+
+# Select columns before column J
+cleaned_data <- raw_data[, 1:(start_index - 1)]
+
+cleaned_data <- janitor::clean_names(cleaned_data)
 
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+write_csv(cleaned_data, "data/analysis_data.csv")

### 2024-04-21 paper/paper.qmd
commit

@@
 library(palmerpenguins)
+
+analysis_data <- read_csv(here::here("data/analysis_data.csv"))
 ```
@@
 
+The data in this post was compiled using open source statistical programming language R (RCoreTeam).
+It uses functionalists from readxl, dplyr, ggplot2, here, and readr.
+The cleaning process focuses on transforming the data from the third column onward.
+It uses the mutate and across functions in dplyr to standardize the responses by converting any values that are not "yes" or "no" to NA (missing values), ensuring that the data set contains only clear binary responses for analysis.
+This step is critical to maintaining the consistency and reliability of the data set, as it filters out ambiguous or invalid entries.
+Finally, the cleaned data is saved as a CSV file in the specified "analysis_data" directory.
+This allows access to the cleaned data set for further analysis or reporting.
+Using write.csv with line name = FALSE ensures the output file is neat.
+
 Data Cleaning
+
+To clean the data for better analysis processing, necessary columns such as how many bedrooms are in the social housing units and repetitive columns were removed.
 
@@
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
 The maps of Toronto showing where those social housing are spread gives the reader a more specific sense of what they look like in Toronto.
@@
 Additionally, it breaks down the waiting list by household type (e.g., no dependents, with dependents, seniors).
+
+**Figure 1:**
 
@@
 # Extract relevant columns with adjusted column names
-year_quarter <- raw_data$Quarter_Year
-new_reactivated_applications <- raw_data$New_reactivated
-housed <- raw_data$Housed
+year_quarter <- analysis_data$quarter_year
+new_reactivated_applications <- analysis_data$new_reactivated
+housed <- analysis_data$housed
 
 # Create a data frame
-plot_data <- data.frame(Year_Quarter = year_quarter,
-                        New_Reactivated_Applications = new_reactivated_applications,
-                        Housed = housed)
+plot_data <- data.frame(
+              quarter_year = year_quarter,
+              test = new_reactivated_applications,
+              housed = housed
+            )
 
-# Convert Year_Quarter to factor to maintain order
-plot_data$Year_Quarter <- factor(plot_data$Year_Quarter, levels = unique(plot_data$Year_Quarter))
+# Convert Quarter/Year to factor to maintain order
+plot_data$quarter_year <- 
+  factor(plot_data$quarter_year, levels = unique(plot_data$quarter_year))
 
-# Create a line plot
-ggplot(plot_data, aes(x = Year_Quarter)) +
-  geom_line(aes(y = New_Reactivated_Applications, color = "New/Reactivated Applications"), size = 1.2) +
-  geom_line(aes(y = Housed, color = "Housed"), linetype = "dashed", size = 1.2) +
+# Create a plot
+ggplot(plot_data, 
+       aes(x = quarter_year)) +
+  geom_point(aes(y = test), color = "red") +
+  geom_point(aes(y = housed), color = "blue") +
   scale_color_manual(values = c("blue", "red")) +
@@
 
+**Figure 2:**
+
 ```{r}
@@
 # Extract relevant columns with adjusted column names
-quarter/year <- raw_data$Quarter/Year
-total_active <- raw_data$Total_active_waiting_list
-seniors <- raw_data$Seniors
+year_quarter <- analysis_data$quarter_year
+new_reactivated_applications <- analysis_data$new_reactivated
+housed <- analysis_data$housed
 
 # Create a data frame
-plot_data <- data.frame(Quarter/Year = quarter/year,
-                        Total_active_waiting_list = total_active,
-                        Seniors = seniors)
+plot_data <- data.frame(
+              quarter_year = year_quarter,
+              test = new_reactivated_applications,
+              housed = housed
+            )
 
 # Convert Quarter/Year to factor to maintain order
-plot_data$Quarter/Year <- factor(plot_data$Quarter/Year, levels = unique(plot_data$Year/Quarter))
+plot_data$quarter_year <- 
+  factor(plot_data$quarter_year, levels = unique(plot_data$quarter_year))
 
 # Create a line plot
-ggplot(plot_data, aes(x = Quarter_Year)) +
+ggplot(plot_data, aes(x = quarter_year)) +
   geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Applications"), size = 1.2) +
@@
 Overall, this graph underscores the importance of considering demographic diversity and changing trends in social housing applications when developing and implementing policies and programs aimed at addressing housing needs and promoting housing equity for all segments of the population.
+
+**Figure 3:**
 
@@
 Such insights can inform evidence-based policy interventions and targeted strategies aimed at addressing housing disparities and promoting equitable access to affordable and suitable housing for all households, regardless of their composition.
+
+**Figure 4:**
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Load required packages
+library(ggplot2)
+
+# Extract relevant columns with adjusted column names
+year_quarter <- analysis_data$quarter_year
+new_reactivated_applications <- analysis_data$new_reactivated
+housed <- analysis_data$housed
+
+# Create a data frame
+plot_data <- data.frame(
+              quarter_year = year_quarter,
+              test = new_reactivated_applications,
+              housed = housed
+            )
+
+# Convert Quarter/Year to factor to maintain order
+plot_data$quarter_year <- 
+  factor(plot_data$quarter_year, levels = unique(plot_data$quarter_year))
+
+# Extract relevant columns
+inactive_cancelled <- analysis_data$inactive_cancelled
+total_active_waiting_list <- analysis_data$total_active_waiting_list
+
+# Add the new data to the existing plot_data data frame
+plot_data$Inactive_Cancelled <- inactive_cancelled
+plot_data$Total_Active_Waiting_List <- total_active_waiting_list
+
+# Create a line plot
+ggplot(plot_data, aes(x = quarter_year)) +
+  geom_line(aes(y = Inactive_Cancelled, color = "Inactive/Cancelled"), size = 1.2) +
+  geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Waiting List"), linetype = "dashed", size = 1.2) +
+  scale_color_manual(values = c("blue", "red")) +
+  labs(title = "Inactive/Cancelled vs Total Active Waiting List",
+       x = "Quarter Year",
+       y = "Application Status") +
+  theme_minimal() +
+  theme(legend.title = element_blank())
+
+```
 

### 2024-04-21 paper/paper.qmd
commit

@@
 
-Data is sourced from Open Data Toronto.
+This data is sourced from Open Data Toronto, a public platform that offers a wide array of datasets relevant to various aspects of urban life, including housing, transportation, public health, and more.
+This initiative allows researchers, policymakers, and the general public to access and utilize city-managed data in an effort to foster transparency, drive innovation, and facilitate informed decision-making.
+Utilizing such open data platforms ensures that the information is not only readily available but also consistently updated and maintained according to government standards.
+This approach helps in compiling comprehensive and reliable datasets that are crucial for conducting thorough analyses, like the study of housing application trends, thereby contributing to more effective urban planning and policy development.
 
-The data in this post was compiled using open source statistical programming language R (RCoreTeam).
-It uses functionalists from readxl, dplyr, ggplot2, here, and readr.
-The cleaning process focuses on transforming the data from the third column onward.
-It uses the mutate and across functions in dplyr to standardize the responses by converting any values that are not "yes" or "no" to NA (missing values), ensuring that the data set contains only clear binary responses for analysis.
-This step is critical to maintaining the consistency and reliability of the data set, as it filters out ambiguous or invalid entries.
+The data in this post was compiled using open source statistical programming language R @RCoreTeam.
+It uses functionalists such as janitor and ggplot2.
 Finally, the cleaned data is saved as a CSV file in the specified "analysis_data" directory.
@@
 
-To clean the data for better analysis processing, necessary columns such as how many bedrooms are in the social housing units and repetitive columns were removed.
+To clean the data for better analysis processing, unnecessary columns such as how many bedrooms are in the social housing units and repetitive columns were removed.
+We use the function 'janitor' to achieve intended data cleaning.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: applications
+#| fig-cap: New and total applications comparison
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# Load the required libraries
+library(dplyr)
+
+# First, ensure that the Quarter/Year column is read correctly
+data$Quarter_Year <- as.character(analysis_data$Quarter_Year)
+
+# Filter data for specific quarters
+specific_quarters <- data %>%
+  filter(Quarter_Year %in% c("Q1 2019", "Q1 2020", "Q1 2022", "Q1 2023")) %>%
+  select(Quarter_Year, New_reactivated = New_reactivated, Total_active_waiting_list = Total_active_waiting_list)
+
+# View the table
+specific_quarters
+
 ```
@@
 ggplot(plot_data, aes(x = quarter_year)) +
-  geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Applications"), size = 1.2) +
+  geom_line(aes(y = total_active_waiting_list, color = "Total Active Applications"), size = 1.2) +
   geom_line(aes(y = Seniors, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
@@
 
-Weaknesses and next steps should also be included.
+One potential weakness in the data and its processing pertains to the level of details and completeness of the data collected.
+The dataset only provides aggregated numbers for new/reactivated applications, housed individuals, inactive/cancelled applications, and the total active waiting list, segmented by quarter and year.
+Such aggregation may obscure underlying patterns or anomalies that could be important for more detailed analysis.
+For instance, without data on the reasons for application cancellations or the demographic details of applicants, it's difficult to pinpoint specific challenges or needs within the housing system.
+Furthermore, if the dataset lacks consistent data collection methods across quarters or contains missing data for certain quarters, this could lead to skewed analyses and unreliable conclusions.
+
+Regarding the processing and visualization aspect in R, while the script effectively generates a comparative line plot, it does not incorporate any statistical analysis to assess trends or significant changes over time.
+The choice to display only two metrics (inactive/cancelled and total active waiting list) might oversimplify the complex dynamics of housing needs and availability.
+Additionally, the plot assumes that the viewer will understand the implications of the data trends without providing contextual aids such as confidence intervals, which could inform the viewer about the variability or certainty of the plotted trends.
+The visualization might benefit from a multi-dimensional approach, perhaps including other relevant metrics or employing different types of visualizations to provide a more comprehensive view of the data.
+This would allow stakeholders to make more informed decisions based on a clearer understanding of all influencing factors.
+To enhance our understanding and analysis of the housing data, the next steps should involve increasing the detail level and expanding the scope of data collected.
+This could include gathering demographic information about applicants, such as age, income, and family size, as well as reasons for application cancellations or reactivations.
+Implementing more advanced statistical techniques to identify trends and correlations over time would also be beneficial.
+Additionally, integrating geographic data could help identify regional disparities and specific areas of need.
+These efforts would not only provide a more comprehensive view of the housing situation but also aid in tailoring housing policies and interventions to better meet the specific needs of different populations.
 

### 2024-04-21 paper/paper.qmd
commit

@@
 
-# Create a line plot
+# Create a plot
 ggplot(plot_data, aes(x = quarter_year)) +
-  geom_line(aes(y = total_active_waiting_list, color = "Total Active Applications"), size = 1.2) +
+  geom_line(aes(y = test, color = "Total Active Applications"), size = 1.2) +
   geom_line(aes(y = Seniors, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
@@
 
-# Assuming your dataset is loaded into a variable named 'raw_data'
 # Extract relevant columns with adjusted column names
-year_quarter <- raw_data$Quarter_Year
-household_no_dependents <- raw_data$Household_no_dependents
-household_with_dependents <- raw_data$Household_with_dependents
+year_quarter <- analysis_data$Quarter_Year
+household_no_dependents <- analysis_data$Household_no_dependents
+household_with_dependents <- analysis_data$Household_with_dependents
 

### 2024-04-21 paper/paper.qmd
commit

@@
   geom_line(aes(y = test, color = "Total Active Applications"), size = 1.2) +
-  geom_line(aes(y = Seniors, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
+  geom_line(aes(y = test, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
   scale_color_manual(values = c("blue", "red")) +

### 2024-04-21 paper/paper.qmd
commit

@@
 ```
-The line graph illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
+The dot graph illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
 The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020.
@@
 ```
-The line plot comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
+The dot plot comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
 From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period.

### 2024-04-21 paper/paper.qmd
commit

@@
 plot_data <- data.frame(Year_Quarter = year_quarter,
-                        Household_no_dependents = household_no_dependents,
-                        Household_with_dependents = household_with_dependents)
+                        household_no_dependents = household_no_dependents,
+                        household_with_dependents = household_with_dependents)
 
@@
 ggplot(plot_data, aes(x = Year_Quarter)) +
-  geom_line(aes(y = Household_no_dependents, color = "Household without Dependents"), size = 1.2) +
-  geom_line(aes(y = Household_with_dependents, color = "Household with Dependents"), linetype = "dashed", size = 1.2) +
+  geom_line(aes(y = household_no_dependents, color = "Household without Dependents"), size = 1.2) +
+  geom_line(aes(y = household_with_dependents, color = "Household with Dependents"), linetype = "dashed", size = 1.2) +
   scale_color_manual(values = c("blue", "red")) +

### 2024-04-21 paper/paper.qmd
commit

@@
 ```
-The dot graph illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
+The dot graph @figure1 illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
 The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020.
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
+#| message: false
 
@@
 ```
-The dot plot comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
+The dot plot @figure2 comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
 From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period.
@@
 ggplot(plot_data, aes(x = Year_Quarter)) +
-  geom_line(aes(y = household_no_dependents, color = "Household without Dependents"), size = 1.2) +
-  geom_line(aes(y = household_with_dependents, color = "Household with Dependents"), linetype = "dashed", size = 1.2) +
+  geom_point(aes(y = household_no_dependents, color = "Household without Dependents"), size = 1.2) +
+  geom_point(aes(y = household_with_dependents, color = "Household with Dependents"), linetype = "dashed", size = 1.2) +
   scale_color_manual(values = c("blue", "red")) +
@@
 
-The line plot comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
+The line plot @figure3 comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
 From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period.
@@
 ggplot(plot_data, aes(x = quarter_year)) +
-  geom_line(aes(y = Inactive_Cancelled, color = "Inactive/Cancelled"), size = 1.2) +
-  geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Waiting List"), linetype = "dashed", size = 1.2) +
+  geom_point(aes(y = Inactive_Cancelled, color = "Inactive/Cancelled"), size = 1.2) +
+  geom_point(aes(y = Total_Active_Waiting_List, color = "Total Active Waiting List"), linetype = "dashed", size = 1.2) +
   scale_color_manual(values = c("blue", "red")) +

### 2024-04-21 paper/paper.qmd
commit

@@
 ggplot(plot_data, aes(x = quarter_year)) +
-  geom_line(aes(y = test, color = "Total Active Applications"), size = 1.2) +
-  geom_line(aes(y = test, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
+  geom_point(aes(y = test), color = "red") +
+  geom_point(aes(y = housed), color = "blue") +
   scale_color_manual(values = c("blue", "red")) +

### 2024-04-22 paper/paper.qmd
commit

@@
 
+**Figure 1:**
+
 ```{r}
 #| label: applications
-#| fig-cap: New and total applications comparison
+#| fig-cap: Data Overview
 #| echo: false
 
-# Load the required libraries
-library(dplyr)
+library(knitr)
 
-# First, ensure that the Quarter/Year column is read correctly
-data$Quarter_Year <- as.character(analysis_data$Quarter_Year)
+display<-select(analysis_data,quarter_year, household_no_dependents, household_with_dependents, new_reactivated,total_active_waiting_list)
 
-# Filter data for specific quarters
-specific_quarters <- data %>%
-  filter(Quarter_Year %in% c("Q1 2019", "Q1 2020", "Q1 2022", "Q1 2023")) %>%
-  select(Quarter_Year, New_reactivated = New_reactivated, Total_active_waiting_list = Total_active_waiting_list)
-
-# View the table
-specific_quarters
+kable(head(analysis_data))
 
@@
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-The maps of Toronto showing where those social housing are spread gives the reader a more specific sense of what they look like in Toronto.
+Above @figure1 is an overview of the data. It simulates the first 6 lines of all columns to give the readers an idea of what the data looks like.
 
@@
 
-**Figure 1:**
+**Figure 2:**
 
@@
 ```
-The dot graph @figure1 illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
+The dot graph @figure2 illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
 The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020.
@@
 
-**Figure 2:**
+**Figure 3:**
 
@@
 year_quarter <- analysis_data$quarter_year
-new_reactivated_applications <- analysis_data$new_reactivated
-housed <- analysis_data$housed
+test <- analysis_data$total_active_waiting_list
+seniors <- analysis_data$seniors
 
@@
               quarter_year = year_quarter,
-              test = new_reactivated_applications,
-              housed = housed
+              test = total_active_waiting_list,
+              seniors = seniors
             )
@@
 ```
-The dot plot @figure2 comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
+The dot plot @figure3 comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
 From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period.
@@
 
-**Figure 3:**
+**Figure 4:**
 
@@
 ggplot(plot_data, aes(x = Year_Quarter)) +
-  geom_point(aes(y = household_no_dependents, color = "Household without Dependents"), size = 1.2) +
-  geom_point(aes(y = household_with_dependents, color = "Household with Dependents"), linetype = "dashed", size = 1.2) +
+  geom_point(aes(y = household_no_dependents, color = "red")) +
+  geom_point(aes(y = household_with_dependents, color = "blue")) +
   scale_color_manual(values = c("blue", "red")) +
@@
 
-The line plot @figure3 comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
+The line plot @figure4 comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
 From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period.
@@
 
-**Figure 4:**
+**Figure 5:**
 
@@
               quarter_year = year_quarter,
-              test = new_reactivated_applications,
-              housed = housed
+              test = total_active_waiting_list ,
+              inactive_cancelled = inactive_cancelled
             )
@@
 ggplot(plot_data, aes(x = quarter_year)) +
-  geom_point(aes(y = Inactive_Cancelled, color = "Inactive/Cancelled"), size = 1.2) +
-  geom_point(aes(y = Total_Active_Waiting_List, color = "Total Active Waiting List"), linetype = "dashed", size = 1.2) +
+  geom_point(aes(y = Inactive_Cancelled, color = "Inactive/Cancelled")) +
+  geom_point(aes(y = Total_Active_Waiting_List, color = "Total Active Waiting List")) +
   scale_color_manual(values = c("blue", "red")) +
@@
 ```
+The dot plot @figure5 visualizes the relationship between inactive/cancelled applications and the total applications on the waiting list in Toronto's social housing sector from Q1 2019 through Q4 2023.
+The graph demonstrates the fluctuating nature of inactive or cancelled applications across various quarters, with notable spikes and declines that provide insights into the dynamics affecting the housing waiting list.
+For example, a substantial rise in cancellations or inactive statuses during specific periods could indicate shifts in applicant circumstances or dissatisfaction with the waiting times, which in turn impacts the overall efficiency of housing allocation.
+These fluctuations highlight the challenges in maintaining an active and accurate waiting list, emphasizing the need for robust management systems to handle the complexities of application statuses.
+
+Analyzing the total active waiting list alongside inactive or cancelled applications further illuminates the challenges faced by the housing sector in managing demand.
+The overall trend of the waiting list, when viewed in conjunction with the inactive or cancelled applications, suggests periods of increased strain on housing resources, possibly driven by economic factors or changes in policy.
+Such analysis is critical for housing authorities as it points out the need for adaptive strategies that can swiftly respond to changes in demand and applicant status.
+This insight is particularly important for planning and implementing policies aimed at reducing waiting times and improving the responsiveness of housing services to better meet the needs of those on the waiting list.
 
@@
 This would allow stakeholders to make more informed decisions based on a clearer understanding of all influencing factors.
+
 To enhance our understanding and analysis of the housing data, the next steps should involve increasing the detail level and expanding the scope of data collected.

### 2024-04-22 paper/paper.qmd
commit

@@
 year_quarter <- analysis_data$Quarter_Year
-household_no_dependents <- analysis_data$Household_no_dependents
-household_with_dependents <- analysis_data$Household_with_dependents
+tes1 <- analysis_data$Household_no_dependents
+test2 <- analysis_data$Household_with_dependents
 
@@
 plot_data <- data.frame(Year_Quarter = year_quarter,
-                        household_no_dependents = household_no_dependents,
-                        household_with_dependents = household_with_dependents)
+                        test1 = household_no_dependents,
+                        test2 = household_with_dependents)
 

### 2024-04-22 paper/paper.qmd
commit

@@
 From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period.
-However, seniors' applications appear to comprise a relatively smaller proportion of the total active applications, as indicated by the consistently lower values for seniors' applications compared to total active applications.
+However, seniors' applications appear to comprise a consistantly smaller proportion of the total active applications, as indicated by the consistently lower values for seniors' applications compared to total active applications.
 This disparity suggests that while seniors constitute a significant demographic group in need of social housing support, they may not represent the majority of applicants.
@@
 The graph demonstrates the fluctuating nature of inactive or cancelled applications across various quarters, with notable spikes and declines that provide insights into the dynamics affecting the housing waiting list.
-For example, a substantial rise in cancellations or inactive statuses during specific periods could indicate shifts in applicant circumstances or dissatisfaction with the waiting times, which in turn impacts the overall efficiency of housing allocation.
+For example, a substantial rise in cancellations or inactive statuses during specific periods, which are quarter 2 2019 and quarter 1 2021, could indicate shifts in applicant circumstances or dissatisfaction with the waiting times, which in turn impacts the overall efficiency of housing allocation.
 These fluctuations highlight the challenges in maintaining an active and accurate waiting list, emphasizing the need for robust management systems to handle the complexities of application statuses.

### 2024-04-22 paper/paper.qmd
commit

@@
 
-# Data {#sec-data}
-
-Data Sourcing
+# Data Sourcing
 
@@
 
-Data Cleaning
+#Data Cleaning
 
@@
 
-# Create a plot
+# Create a dot plot
 ggplot(plot_data, 
@@
               quarter_year = year_quarter,
-              test = total_active_waiting_list,
+              test = test,
               seniors = seniors
@@
 
-# Create a plot
+# Create a dot plot
 ggplot(plot_data, aes(x = quarter_year)) +
@@
 # Extract relevant columns with adjusted column names
-year_quarter <- analysis_data$Quarter_Year
-tes1 <- analysis_data$Household_no_dependents
-test2 <- analysis_data$Household_with_dependents
+year_quarter <- analysis_data$quarter_year
+test1 <- analysis_data$household_no_dependents
+test2 <- analysis_data$household_with_dependents
 
@@
 plot_data <- data.frame(Year_Quarter = year_quarter,
-                        test1 = household_no_dependents,
-                        test2 = household_with_dependents)
-
-# Convert Year_Quarter to factor to maintain order
-plot_data$Year_Quarter <- factor(plot_data$Year_Quarter, levels = unique(plot_data$Year_Quarter))
+                        test1 = test1,
+                        test2 = test2)
 
@@
 ggplot(plot_data, aes(x = Year_Quarter)) +
-  geom_point(aes(y = household_no_dependents, color = "red")) +
-  geom_point(aes(y = household_with_dependents, color = "blue")) +
+  geom_point(aes(y = test1, color = "red",group = 1)) +
+  geom_point(aes(y = test2, color = "blue",group = 2)) +
   scale_color_manual(values = c("blue", "red")) +
@@
        y = "Application Numbers") +
-  theme_minimal() +
-  theme(legend.title = element_blank())
+  theme_minimal()
 
@@
 
-The line plot @figure4 comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
+The dot plot @figure4 comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
 From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period.
@@
 
-# Create a line plot
+# Create a dot plot
 ggplot(plot_data, aes(x = quarter_year)) +
@@
 
-# Additional data details
-
 # Model details {#sec-model-details}

### 2024-04-22 paper/paper.qmd
commit

@@
 
-Above @figure1 is an overview of the data. It simulates the first 6 lines of all columns to give the readers an idea of what the data looks like.
+The above @figure1 is an overview of the data.
+It simulates the first 6 lines of all columns to give the readers an idea of what the data looks like.
+The complete data consists of 9 columns/variables and 20 lines/quarters of years from quarter 1 2019 to quarter 4 2023.
 
@@
 
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-04-22 paper/paper.qmd
commit

@@
 
-To clean the data for better analysis processing, unnecessary columns such as how many bedrooms are in the social housing units and repetitive columns were removed.
+To clean the data for better analysis processing, unnecessary columns such as how many bedrooms are in the social housing units and the repetitive column 'total application' were removed.
 We use the function 'janitor' to achieve intended data cleaning.
@@
 ```{r}
+#| label: data-overview
+#| #| fig-cap: Data Overview
+#| fig.pos: H
 #| label: applications
@@
 
-The above @figure1 is an overview of the data.
+The above @data-overview is an overview of the data.
 It simulates the first 6 lines of all columns to give the readers an idea of what the data looks like.
@@
 ```{r}
+#| label: new-and-housed
+#| fig.pos: H
 #| echo: false
@@
 ```
-The dot graph @figure2 illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
+The dot plot @new-and-housed illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
 The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020.
@@
 ```{r}
+#| label: total-and-seniors
+#| fig.pos: H
 #| echo: false
@@
 ```
-The dot plot @figure3 comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
+The dot plot @total-and-seniors comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
 From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period.
@@
 ```{r}
+#| label: dependents-or-not
+#| fig.pos: H
 #| echo: false
@@
 
-The dot plot @figure4 comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
+The dot plot @dependents-or-not comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
 From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period.
@@
 ```{r}
+#| label: cancelled-vs-total
+#| fig.pos: H
 #| echo: false
@@
 ```
-The dot plot @figure5 visualizes the relationship between inactive/cancelled applications and the total applications on the waiting list in Toronto's social housing sector from Q1 2019 through Q4 2023.
+The dot plot @cancelled-vs-total visualizes the relationship between inactive/cancelled applications and the total applications on the waiting list in Toronto's social housing sector from Q1 2019 through Q4 2023.
 The graph demonstrates the fluctuating nature of inactive or cancelled applications across various quarters, with notable spikes and declines that provide insights into the dynamics affecting the housing waiting list.
@@
 
-\newpage
+# Conclusion
+This study has provided critical insights into the complex dynamics of household composition within Toronto's social housing applications from Q1 2019 to Q4 2023.
+By closely analyzing the data trends, the research highlights the significant challenges that families, particularly those with dependents, face in accessing stable and affordable housing.
+The findings underscore a marked inconsistency between the demand for social housing and the availability of accommodations, reflecting a substantial gap that persists despite varying applications rates.
+This persistent discrepancy indicates that current housing policies and resources may not adequately address the needs of the most vulnerable populations.
+As such, the study calls for a deeper and more nuanced understanding of these trends to facilitate the development of targeted policy interventions aimed at promoting housing equity and stability for families in Toronto.
+These efforts are crucial for not only improving living conditions but also for ensuring that housing policies evolve in alignment with the changing demographics and economic realities of the community.
 
-\appendix
-
-# Appendix {-}
+Moving forward, the paper suggests several avenues for further research and policy development.
+It emphasizes the importance of enhancing data collection processes to capture more granular, demographic-specific information that can inform more effective housing strategies.
+Additionally, there is a strong recommendation for integrating adaptive policies that can swiftly respond to the fluctuations in housing demand and the specific circumstances of applicants.
+Such policies should focus on reducing waiting times, optimizing resource allocation, and improving the overall responsiveness of housing services.
+By addressing these critical issues, Toronto can make significant strides towards reducing housing inequities and ensuring that all residents have access to adequate and affordable housing.
+This comprehensive approach will not only address the immediate needs but also contribute to the long-term sustainability and resilience of the city’s housing system.
 

### 2024-04-22 paper/paper.qmd
commit

@@
 #| label: data-overview
-#| #| fig-cap: Data Overview
-#| fig.pos: H
-#| label: applications
 #| fig-cap: Data Overview
@@
 year_quarter <- analysis_data$quarter_year
-new_reactivated_applications <- analysis_data$new_reactivated
-housed <- analysis_data$housed
+test <- analysis_data$total_active_waiting_list
+inactive_cancelled <- analysis_data$inactive_cancelled
 

### 2024-04-22 paper/paper.qmd
commit

@@
 library(palmerpenguins)
+library(tinytex)
 
 analysis_data <- read_csv(here::here("data/analysis_data.csv"))
+
 ```
@@
               quarter_year = year_quarter,
-              test = total_active_waiting_list ,
+              test = test ,
               inactive_cancelled = inactive_cancelled

### 2024-04-22 scripts/03-test_data.R

@@ -1,15 +1,24 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
+# Purpose: Tests 
+# Author: Xiaoyu Miao
 # Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Contact: xy.miao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Have the data downloaded
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Test data ####
+
+# Load data
+data <- read.csv(here::here("data/analysis_data.csv"))
+
+# Test to ensure all values are greater than zero
+test_results <- data %>% 
+  summarise(across(everything(), ~ all(. > 0)))
+
+# Print the test results
+print(test_results)
+

### 2024-04-22 paper/paper.qmd
commit

@@
 ```{r}
-#| label: data-overview
-#| fig-cap: Data Overview
+#| label: tbl-data-overview
+#| tbl-cap: Data Overview
 #| echo: false
@@
 
-The above @data-overview is an overview of the data.
+The above @tbl-data-overview is an overview of the data.
 It simulates the first 6 lines of all columns to give the readers an idea of what the data looks like.
@@
 ```{r}
-#| label: new-and-housed
+#| label: fig-new-and-housed
 #| fig.pos: H
@@
 ```
-The dot plot @new-and-housed illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
-The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020.
-Despite these fluctuations, the number of households housed each quarter consistently lags behind new entries, underscoring a substantial gap between demand and supply dynamics.
-This persistent shortfall indicates the ongoing challenge of meeting the housing needs of vulnerable populations within the community.
+The dot plot @fig-new-and-housed illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023.
+The red dots signify the number of yearly new or reactivated applications for social housing, while the blue dots illustrate the number of households that were successfully housed each year.
+A striking feature of the plot is the consistent positioning of the red dots well above the blue dots, which clearly indicates that every year, the number of new or reactivated applications far exceeds the number of households that are actually housed.
 
-Examining waiting list trends further illuminates the complexities involved in managing social housing demand.
-The total active waiting list experiences fluctuations, notably declining in Q2 2019 due to a significant number of inactive or cancelled applications.
-This fluctuation suggests challenges in maintaining accurate waiting lists or changes in applicants' circumstances, highlighting the multifaceted nature of addressing social housing needs.
-Such insights underscore the importance of robust data management and adaptive policies to effectively respond to evolving housing demands.
+A specific observation is that the number of households housed—indicated by the blue dots—remains relatively low and consistent over time, suggesting a steady capacity or rate at which the housing sector is able to accommodate applicants.
+In contrast, the number of new or reactivated applications represented by the red dots shows more variability, with certain years experiencing higher peaks.
+This variability could reflect economic cycles, changes in housing policy, or other societal factors that influence the number of individuals seeking social housing.
+The peaks could correspond to economic downturns or policy changes that either increase the availability of social housing or change the criteria for application, resulting in more people applying.
 
-Moreover, an analysis of the waiting list composition reveals critical insights into the demographics of those in need of social housing.
-A significant portion of seniors and households without dependents feature prominently in the waiting list, indicating a pressing need for housing support among these vulnerable groups.
-This demographic composition underscores broader social and economic challenges, including affordability issues for seniors and single individuals, necessitating targeted interventions and policy measures to address housing inequities and ensure inclusive access to adequate housing for all segments of the population.
+The consistent gap between the red and blue dots underscores a significant challenge within the social housing system: the demand for housing assistance consistently outstrips supply.
+This suggests that despite the efforts to house individuals and families in need, there remains a substantial number of applicants who are left on the waiting list, potentially leading to prolonged periods of housing insecurity and uncertainty for a considerable number of residents.
+For policymakers and social planners, this graph is a call to action, indicating the need to increase the availability of social housing, streamline the application process, or explore alternative housing solutions to bridge the gap between application and accommodation.
+The social implications are profound, as housing stability is closely tied to other areas of well-being, including employment, education, and mental health.
+Therefore, the insights from this graph are vital for informing a more strategic and compassionate response to housing needs in Toronto.
 
@@
 ```{r}
-#| label: total-and-seniors
+#| label: fig-total-and-seniors
 #| fig.pos: H
@@
 ```
-The dot plot @total-and-seniors comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
-From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period.
-However, seniors' applications appear to comprise a consistantly smaller proportion of the total active applications, as indicated by the consistently lower values for seniors' applications compared to total active applications.
-This disparity suggests that while seniors constitute a significant demographic group in need of social housing support, they may not represent the majority of applicants.
-Understanding this dynamic is crucial for policymakers and housing providers to tailor their strategies and resources effectively to address the diverse needs of applicants across different age groups.
+The dot plot @fig-total-and-seniors comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time.
+The red dots at the top of the graph represent the total active applications, which appear to hold steady at a high level throughout the entire period.
+In contrast, the blue dots at the bottom of the graph depict seniors' applications for social housing, which are significantly lower in number compared to the total applications and remain consistent over time.
 
-Moreover, the fluctuating trends in both total active applications and seniors' applications highlight the dynamic nature of housing demand and the complex socio-economic factors influencing individuals' housing situations.
-Further analysis may be warranted to explore the underlying reasons behind these fluctuations, such as changes in demographic trends, economic conditions, or shifts in housing policies and programs.
-Overall, this graph underscores the importance of considering demographic diversity and changing trends in social housing applications when developing and implementing policies and programs aimed at addressing housing needs and promoting housing equity for all segments of the population.
+Specifically, the graph shows a persistent gap between the total number of applications and the number attributed to seniors, indicating that seniors are a smaller, consistent group within the overall pool of applicants.
+The fact that the number of seniors' applications is much lower than the total might suggest that the social housing system is either not as utilized by seniors or that there are not enough seniors applying or qualifying for social housing within the reporting periods.
+This could point to a need for more targeted outreach or specialized housing policies that cater to the elderly population's specific requirements.
+
+The high and stable total application numbers highlight the ongoing demand for social housing and suggest that this demand is not significantly driven by seniors but by other demographic groups.
+The constancy of the total applications suggests that the issue of housing insecurity is widespread and persists over time, reflecting systemic issues within the housing market or broader socio-economic challenges in Toronto.
+For social policymakers, these consistent figures could indicate a need for sustained or increased investment in social housing to keep pace with demand and address the long-term needs of the city's diverse population.
 
@@
 ```{r}
-#| label: dependents-or-not
+#| label: fig-dependents-or-not
 #| fig.pos: H
@@
 
-The dot plot @dependents-or-not comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
+The dot plot @fig-dependents-or-not comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time.
+Red dots, representing households with dependents, show more pronounced fluctuations when compared to blue dots, which symbolize households without dependents.
+This suggests a higher degree of instability or variability in housing needs among families, which could be a reflection of the social and economic volatility that impacts families more intensely than individuals.
+The red dots peak and trough more dramatically, signaling times when families are perhaps feeling the economic squeeze—times when social support becomes even more critical.
 From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period.
@@
 
-Moreover, further analysis of the underlying factors driving fluctuations in applications from households with and without dependents may provide valuable insights into broader socio-economic trends and challenges affecting housing affordability, access, and stability for families in need.
+Furthermore, the steadier trend seen with individuals without dependents could point to different social implications.
+While perhaps not as acutely affected by economic shifts, this group still requires attention to ensure that the available social housing meets their needs, which may differ significantly from those of families.
+This data prompts a deeper consideration of how social housing is allocated and managed, suggesting a need for programs that are sensitive to the nuances of household composition.
+Policies must be nuanced and empathetic, ensuring equity not just in housing provision but in the opportunity for all citizens to live with dignity and security.
+Such social housing strategies become foundational to building a resilient society where every individual, family, or household can weather economic storms without the risk of homelessness or displacement.
+Further analysis of the underlying factors driving fluctuations in applications from households with and without dependents may provide valuable insights into broader socio-economic trends and challenges affecting housing affordability, access, and stability for families in need.
 Such insights can inform evidence-based policy interventions and targeted strategies aimed at addressing housing disparities and promoting equitable access to affordable and suitable housing for all households, regardless of their composition.
@@
 ```{r}
-#| label: cancelled-vs-total
+#| label: fig-cancelled-vs-total
 #| fig.pos: H
@@
 ```
-The dot plot @cancelled-vs-total visualizes the relationship between inactive/cancelled applications and the total applications on the waiting list in Toronto's social housing sector from Q1 2019 through Q4 2023.
-The graph demonstrates the fluctuating nature of inactive or cancelled applications across various quarters, with notable spikes and declines that provide insights into the dynamics affecting the housing waiting list.
-For example, a substantial rise in cancellations or inactive statuses during specific periods, which are quarter 2 2019 and quarter 1 2021, could indicate shifts in applicant circumstances or dissatisfaction with the waiting times, which in turn impacts the overall efficiency of housing allocation.
-These fluctuations highlight the challenges in maintaining an active and accurate waiting list, emphasizing the need for robust management systems to handle the complexities of application statuses.
+The dot plot @fig-cancelled-vs-total visualizes the relationship between inactive/cancelled applications and the total applications on the waiting list in Toronto's social housing sector from Q1 2019 through Q4 2023.
+This dot plot visualizes two data points related to social housing applications over several quarters from Q1 2019 to Q4 2023.
+The red dots represent the total number of active applications on the waiting list, while the blue dots signify the number of applications that have been either canceled or marked as inactive.
+Observing the graph specifically, the red dots are positioned at a high level consistently across the timeline, indicating a stable and significant demand for social housing, as the active waiting list numbers are quite high and do not show dramatic changes over time.
+On the other hand, the blue dots are much lower on the graph, which shows that the number of inactive or canceled applications is significantly less than the total number of active applications.
+However, despite being lower, the blue dots are not negligible, suggesting that there is still a considerable number of applications that are removed from the waiting list for various reasons.
+This is especially the case in Q2 2019 and Q1 2021, where a significantly larger-than-average numbers of applications were cancelled or became inactive.
 
-Analyzing the total active waiting list alongside inactive or cancelled applications further illuminates the challenges faced by the housing sector in managing demand.
-The overall trend of the waiting list, when viewed in conjunction with the inactive or cancelled applications, suggests periods of increased strain on housing resources, possibly driven by economic factors or changes in policy.
-Such analysis is critical for housing authorities as it points out the need for adaptive strategies that can swiftly respond to changes in demand and applicant status.
-This insight is particularly important for planning and implementing policies aimed at reducing waiting times and improving the responsiveness of housing services to better meet the needs of those on the waiting list.
+The distance between the red and blue dots suggests a discrepancy between the number of people seeking housing and those who leave the system without being housed.
+This gap may highlight inefficiencies or barriers within the social housing system that result in applicants dropping out or being removed from the list.
+The stability in the number of active applications suggests a persistent need for social housing, while the presence of inactive or canceled applications could point to potential improvements in the system to better retain and service applicants.
+
+From a policy perspective, this graph calls attention to the potential for improving the housing allocation process to decrease the number of inactive/canceled applications and better meet the housing needs represented by the consistently high active waiting list.
+It could also indicate the need to examine the reasons behind the cancellations and inactivity, as addressing these could lead to higher efficiency within the social housing system and increased satisfaction among applicants.
 

### 2024-04-22 scripts/00-simulate_data.R

@@ -12,15 +12,17 @@ library(tidyverse)
 #### Simulate data ####
 # This setup will create a structured tabular dataset where each row represents the social housing status for a given year and applications ID, and the active applications are recorded in the new/reactive column, the inactive/cancelled applications are recorded in the inactive/cancelled column. This dataset is constructed to reflect different scenarios in the OpenDataToronto where it collects information from quarters of different years. However, columns such as "special priority","over housed" were reduced to keep the data streamlined.
 
-set.seed(1688) # For reproducibility
-years <- 2019:2023 # Create a sequence of years from 1972 to 2022.
-years <- sort(sample(years, length(years), replace = FALSE)) # Randomly sample from these years without replacement, ensuring that each year in the resulting year vector is unique, and then sort in ascending order.
-num_appli <- sample(50:100, 1) # Choose a random value between 50 and 100 to determine the number of application or records in the modelled data set.
-ids <- 1:num_appli # Generate a string of numbers from 1 to num_appli as a unique identifier for each record or application in the dataset.
-statuses <- sample(c("New/reactive", "Housed", "Inactive/cancelled"), num_appli, replace = TRUE) # Statuses are randomly selected from the "New/reactive", "Housed" and "Inactive/Cancelled" options in num_appli, allowing for substitution, which means that the same status can be selected more than once for different applications.
-
-data <- data.frame(year = rep(years, length.out = num_appli),# Repeats the year vector to match the total number of application. If num_appli is greater than the length of the year, it loops through the year vector until num_appli is reached.
-                   id = ids,
-                   application_statuses = applicationstatuses)
-
+simulated_dataset <-
+  tibble(
+    year = 2000 + rep(19:23, each = 4),
+    quarter = rep(1:4,each = 5),
+    new = sample(100:10000, size = 20, replace = TRUE),
+    housed = sample(100:10000, size = 20, replace = TRUE),
+    inactive = sample(100:10000, size = 20, replace = TRUE),
+    household_no_dependents = sample(100:10000, size = 20, replace = TRUE),
+    household_with_dependents = sample(100:10000, size = 20, replace = TRUE),
+    seniors = sample(100:10000, size = 20, replace = TRUE),
+    total = household_no_dependents + household_with_dependents + seniors
+  )
+simulated_dataset
 

### 2024-04-22 scripts/01-download_data.R

@@ -1,26 +1,32 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Open Data Toronto
+# Author: Xiaoyu Miao
+# Date: April 17th
+# Contact: xy.miao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+#Access to GSS Data
+#Step 1: Visit the Open Data Toronto website
+#Visit the Open Data Toronto website: https://open.toronto.ca/
 
-
+#Step 2: Find relevant datasets
+#You can then use the search function on the website to search for wanted dataset and complete the data collection.
 
 #### Save data ####
-# [...UPDATE THIS...]
+
+#Step 4: Download the data
+#Once you have found the relevant dataset, you can start downloading it. Just select the formate you wanted and click the "download data" button.
+
+#Optional
 # change the_raw_data to whatever name you assigned when you downloaded it.
+
 write_csv(the_raw_data, "inputs/data/raw_data.csv") 
 
          

### 2024-04-22 paper/paper.qmd
commit

@@
 #| label: fig-new-and-housed
+#| fig-cap: Yearly New/Reactivated Applications vs. Housed
 #| fig.pos: H
@@
 #| label: fig-total-and-seniors
+#| fig-cap: Total Active Applications vs. Seniors Applications
 #| fig.pos: H
@@
 #| label: fig-dependents-or-not
+#| fig-cap: Households with and without Dependents Applying for Social Housing
 #| fig.pos: H
@@
 #| label: fig-cancelled-vs-total
+#| fig-cap: Inactive/Cancelled vs Total Active Waiting List
 #| fig.pos: H

### 2024-04-22 paper/paper.qmd
commit

@@
 
+The fluctuations observed in the applications signal shifts in socio-economic conditions that may differentially impact seniors.
+Economic downturns, pension adjustments, and healthcare costs are just a few factors that could influence a senior's need for social housing.
+The relative stability in the number of seniors' applications compared to the total may suggest that while seniors are impacted by these factors, their numbers in the social housing system remain proportionately constant.
+Policymakers and housing providers should, therefore, consider economic and health policies in tandem with housing strategies to address the broader spectrum of senior needs.
+These strategies could include not only traditional housing assistance but also innovative solutions like rent subsidies, housing cooperatives, and community living arrangements that promote social inclusion and improve the overall quality of life for seniors.
+The data presented by the dot plot reinforce the necessity for a strategic, data-informed approach to policy-making, ensuring that all citizens, especially the elderly, have access to adequate, affordable, and appropriate housing options.
+
 ## Third discussion point
@@
 For instance, fluctuations in applications from households with dependents may reflect changes in employment opportunities, income levels, or housing affordability, which impact family decision-making regarding housing choices and stability.
+
+From a social perspective, the implications of this data are multifaceted.
+Households with dependents, often grappling with the higher cost of living and childcare, may face more acute housing crises when confronted with economic challenges.
+The dot plot's variability underscores the precariousness of their housing situation, which can have ripple effects throughout the community, affecting children's education, family health, and overall social stability.
+For policymakers, these insights highlight the urgency of creating flexible housing solutions that can adapt to the cyclical economic pressures faced by families.
+It's not just about providing shelter, but about fostering an environment where families can thrive, contributing positively to the social fabric.
 

### 2024-04-22 paper/paper.qmd
commit

@@
   - Xiaoyu (Alice) Miao
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/xyalicemiao/social-housing-toronto/tree/main"
 date: today

### 2024-04-22 scripts/03-test_data.R
commit data testing

@@ -10,7 +10,19 @@
 #### Workspace setup ####
 library(tidyverse)
 
-#### Test data ####
+#### Test 1 ####
+
+# Load data
+data <- read.csv(here::here("data/analysis_data.csv"))
+
+# Test to ensure there are no missing values in any cell of the dataset
+no_missing_values <- !anyNA(data)
+
+# Print the test result
+print(no_missing_values)
+
+
+#### Test 2 ###
 
 # Load data
 data <- read.csv(here::here("data/analysis_data.csv"))
@@ -21,4 +33,34 @@ test_results <- data %>%
 
 # Print the test results
 print(test_results)
+
+
+#### Test 3 ####
+
+# Load data
+data <- read.csv(here::here("data/analysis_data.csv"))
+
+# Test to ensure all numeric values are integers
+test_results <- data %>% 
+  summarise(across(where(is.numeric), ~ all(. == floor(.))))
+
+# Print the test results
+print(test_results)
+
+
+#### Test 4 ####
+
+# Load data
+data <- read.csv(here::here("data/analysis_data.csv"))
+
+# Test to ensure there are no missing values in the dataset
+test_results <- data %>% 
+  summarise(across(everything(), ~ all(complete.cases(.))))
+
+# Print the test results
+print(test_results)
+
+
+
+
 
